Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-02-26 Grassmannian_cells.tex
Remark on equivalences

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\bfi}{\mathbf{i}}
\newcommand{\bfI}{\mathbf{I}}
\newcommand{\bfJ}{\mathbf{J}}
\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\ui}{{\underline i}}
\newcommand{\uj}{{\underline j}}
\newcommand\udim{{\underline{\dim}\, }}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\rep}{\operatorname{rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Ind}{\mathrm{Ind}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\rk}{\mathrm{rk}}
\newcommand{\cQ}{\mathcal{Q}}
\newcommand{\Att}{\operatorname{Att}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\im}{\operatorname{im}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\into}{\hookrightarrow}
\newcommand{\onto}{\to\!\!\!\!\!\to}
\newcommand{\xonto}[1]{\xrightarrow{#1}\!\!\!\!\!\to}
\newcommand{\ses}[3]{\xymatrix@C15pt{0\ar[r] & #1\ar[r] & #2\ar[r] & #3 \ar[r] & 0}}
\newcommand{\sesm}[4]{\xymatrix{0\ar[r] & #1\ar[r] & #2\ar^{#4}[r] & #3 \ar[r] & 0}}
%\newcommand{\sesm}[4]{0\rightarrow #1\rightarrow #2\xrightarrow{#4}#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\newcommand{\spn}{\operatorname{span}}
\newcommand{\supp}{\operatorname{supp}}
\setcounter{MaxMatrixCols}{20}
\newcommand{\vs}{\vspace{0.2cm}}
\hyphenation{endo-functors}

@@

\author{Dylan Rupel}
\author{Thorsten Weist}

@@

We introduce a notion of equivalence of $2$-quivers.
\begin{definition}
  \label{def:2equivalence}
  Let $\tilde Q=(Q,Q_2)$ be a $2$-quiver with a $2$-arrow $V:\Gamma(1)\to\Gamma(2)$ in $Q_2$ such that one of the following conditions is satisfied:
  \begin{enumerate}
    \item $\Gamma(1)$ has precisely one source $p$; 
    \item $\Gamma(2)$ has precisely one sink $q$;
    \item $\Gamma(1)=\{p\}$ and $\Gamma(2)=\{q\}$.
  \end{enumerate}
  Depending on the condition which is satisfied, we define
  \begin{enumerate}
    \item $\tilde Q_p$ as the $2$-quiver obtained from $\tilde Q$ when replacing the 2-arrow $V$ by a $2$-arrow $V_p:\{p\}\to\Gamma(2)$;
    \item $\tilde Q_q$ as the $2$-quiver obtained from $\tilde Q$ when replacing the 2-arrow $V$ by a $2$-arrow $V_q:\Gamma(1)\to \{q\}$;
    \item $\tilde Q_V$ as the $2$-quiver obtained from $\tilde Q$ when replacing the 2-arrow $V$ by a usual arrow $\alpha_V:p\to q$. 
  \end{enumerate}
  This defines a relation on the set of $2$-quivers denoted by $\tilde Q\to \tilde Q_p$, $\tilde Q\to \tilde Q_q$ and $\tilde Q\to \tilde Q_V$ respectively.
  Moreover, it induces an equivalence relation $\sim$ on the set of $2$-quivers when taking the symmetric and transitive closure of this relation. 
\end{definition}

@@

@@

%\begin{remark}\footnote{Insert a remark that we do not work on the universal cover} 
%  Lemma \ref{AR} can be generalized to the cases $m\leq 2$.
%  The only technical problem is that $\tau \tilde P^{(k)}_{m+1}$ is no quotient of $\tilde P_{m,k}$ and we have to replace it by the image of the unique morphism to $\tau \tilde P^{(k)}_{m+1}$ which induces the exact sequence. 
%\end{remark}
In the following, we frequently use the notation and conventions of Section \ref{Lifting}.
The next step is to associate a $2$-quiver $\mathcal Q_m^{[k]}$ with the lifted truncated preprojective representations 
$$\tilde P_m^{[k]}:=\begin{cases}\tilde P_m^{\{n,\ldots,n-k+1\}}\text{ if } m\text{ is even}\\\tilde P_m^{\{1,\ldots,k\}}\text{ if }m\text{ is odd}\end{cases}$$ 
for $m\geq 2$ and $0\leq k\leq n-1$  whose strong successor closed subsets are in one-to-one correspondence with cells of quiver Grassmannians of $\tilde P_m^{[k]}$. Recall that, by Theorem \ref{celldec}, the cells of the quiver Grassmannians attached to $P_m^{[k]}:= G_{K(n)}\tilde P_m^{[k]}$ are in one-to-one correspondence with those attached to the lift $\tilde P_m^{[k]}$. Moreover, by Theorem \ref{thm:truncpp}, the quiver Grassmannians $\Gr_{\tbfe}(P_m^V)$ of arbitrary truncated preprojective representations $P_m^V$ only depend on $\tbfe$ and $\udim P_m^V$. In short, the $2$-quivers $\mathcal Q_m^{[k]}$ can be seen as being attached to any (truncated) preprojective representation of the respective dimension.

For building the mentioned $2$-quivers, we use the recursive construction described in Section \ref{Lifting}. The following remarks are meant to understand the structure of the $2$-quivers. 
%In particular, we choose bases of the $\Ext$-groups which are compatible with the lifts to the universal covering.
%For simplicity and in order to be in shape(?) with Sections \ref{Lifting} and \ref{sec:fibrations}, we work on the universal cover.
\begin{remark}\begin{enumerate}\item It will turn out that every vertex of the quiver $\mathcal Q_m^{[k]}$ uniquely corresponds to a vertex of the universal covering quiver $\widetilde{K(n)}$, but that different vertices of $\cQ_m^{[k]}$ can correspond to the same vertex of $\widetilde{K(n)}$. Also every usual arrow of $\mathcal Q_m^{[k]}$ does correspond to an arrow of $\widetilde{K(n)}$ and thus to an arrow of $K(n)$. In particular, the quiver obtained from $\mathcal Q_m^{[k]}$ when forgetting the $2$-arrows gives rise to a (non-disjoint) union of subquivers of $\widetilde{K(n)}$.

@@

\noindent
where the $2$-arrows are colored red. Note that the vertices $(1,\alpha_i)$ are just the $2$-quivers $\cQ_{1,i}$ corresponding to $\tilde P_{1,i}$. The $2$-quiver $\mathcal Q_2^{[k]}$ is clearly equivalent to the support quiver (\ref{coeff}). In abuse of notation, we also denote the support quiver by $\mathcal Q_{2}^{[k]}$. In order to keep the illustrations and combinatorics simple, we will work with this quiver instead. We also define the translated two quiver $\cQ_{2,i}$ as those obtained from $\cQ_{2}$ by translating all vertices and ($2$-)arrows by $\alpha_i^{-1}$.

>>>>

\renewcommand{\AA}{\mathbb{A}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\kk}{\Bbbk}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\bfi}{\mathbf{i}}
\newcommand{\bfI}{\mathbf{I}}
\newcommand{\bfJ}{\mathbf{J}}

\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}

\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}}

\newcommand{\ui}{{\underline i}}
\newcommand{\uj}{{\underline j}}
\newcommand\udim{{\underline{\dim}\, }}

\newcommand{\into}{\hookrightarrow}
\newcommand{\onto}{\to\!\!\!\!\!\to}
\newcommand{\xonto}[1]{\xrightarrow{#1}\!\!\!\!\!\to}

\newcommand{\Att}{\operatorname{Att}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gr}{\mathrm{Gr}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\im}{\operatorname{im}}
\newcommand{\Ind}{\mathrm{Ind}}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\rep}{\operatorname{rep}}
\newcommand{\rk}{\mathrm{rk}}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\newcommand{\ses}[3]{\xymatrix@C15pt{0\ar[r] & #1\ar[r] & #2\ar[r] & #3 \ar[r] & 0}}
\newcommand{\sesm}[4]{\xymatrix{0\ar[r] & #1\ar[r] & #2\ar^{#4}[r] & #3 \ar[r] & 0}}
\newcommand{\spn}{\operatorname{span}}
\newcommand{\supp}{\operatorname{supp}}
\newcommand{\vs}{\vspace{0.2cm}}

\hyphenation{endo-functors}
\setcounter{MaxMatrixCols}{20}

@@

\author{Dylan Rupel}
\address[Dylan Rupel]{University of Notre Dame, Department of Mathematics, Notre Dame, IN 46556, USA}
\email{drupel@nd.edu}
\author{Thorsten Weist}
\address[Thorsten Weist]{}
\email{weist@uni-wuppertal.de}

@@

We introduce a notion of equivalence of $2$-quivers.
\begin{definition}
  \label{def:2equivalence}
  Let $\tilde Q=(Q,Q_2)$ be a $2$-quiver with a $2$-arrow $V:\Gamma(1)\to\Gamma(2)$ in $Q_2$ such that one of the following conditions is satisfied:
  \begin{enumerate}
    \item $\Gamma(1)$ has precisely one source $p$; 
    \item $\Gamma(2)$ has precisely one sink $q$;
    \item $\Gamma(1)=\{p\}$ and $\Gamma(2)=\{q\}$.
  \end{enumerate}
  Depending on the condition which is satisfied, we define
  \begin{enumerate}
    \item $\tilde Q_p$ as the $2$-quiver obtained from $\tilde Q$ when replacing the $2$-arrow $V$ by a $2$-arrow $V_p:\{p\}\to\Gamma(2)$;
    \item $\tilde Q_q$ as the $2$-quiver obtained from $\tilde Q$ when replacing the $2$-arrow $V$ by a $2$-arrow $V_q:\Gamma(1)\to \{q\}$;
    \item $\tilde Q_V$ as the $2$-quiver obtained from $\tilde Q$ when replacing the $2$-arrow $V$ by a usual arrow $\alpha_V:p\to q$. 
  \end{enumerate}
  This defines a relation on the set of $2$-quivers denoted by $\tilde Q\to \tilde Q_p$, $\tilde Q\to \tilde Q_q$ and $\tilde Q\to \tilde Q_V$ respectively.
  Moreover, it induces an equivalence relation $\sim$ on the set of $2$-quivers when taking the symmetric and transitive closure of this relation. 
\end{definition}

@@

\begin{remark}
  We will usually apply Lemma~\ref{lem:2equivalence} below after performing each of the equivalences from Definition~\ref{def:2equivalence}.
  That is, given a 2-arrow $V:\Gamma(1)\to\Gamma(2)$ for which $\Gamma(1)$ has a unique source $p$ and $\Gamma(2)$ has a unique sink $q$, we get an equivalent 2-quiver by replacing this 2-arrow with a usual arrow $a_V:p\to q$.
\end{remark}

@@

%\begin{remark}\footnote{Insert a remark that we do not work on the universal cover} 
%  Lemma \ref{AR} can be generalized to the cases $m\leq 2$.
%  The only technical problem is that $\tau \tilde P^{(k)}_{m+1}$ is no quotient of $\tilde P_{m,k}$ and we have to replace it by the image of the unique morphism to $\tau \tilde P^{(k)}_{m+1}$ which induces the exact sequence. 
%\end{remark}
In the following, we frequently use the notation and conventions of Section~\ref{Lifting}.
The next step is to associate a $2$-quiver $\mathcal Q_{m+1}^{[k]}$ with the lifted truncated preprojective representations 
\[\tilde P_{m+1}^{[k]}:=
  \begin{cases}
    \tilde P_{m+1}^{\{n,n-1,\ldots,n-k+1\}} & \text{ if $m$ is odd;}\\
    \tilde P_{m+1}^{\{1,2,\ldots,k\}} & \text{ if $m$ is even;}
  \end{cases}\] 
for $m\geq 1$ and $0\leq k\leq n-1$.
When $k=0$ above, we take $\tilde P_{m+1}^{[0]}=\tilde P_{m+1}$.
We aim to construct $\cQ_{m+1}^{[k]}$ so that the strong successor closed subsets are in one-to-one correspondence with cells of quiver Grassmannians of $\tilde P_{m+1}^{[k]}$.

Recall that, by Theorem~\ref{celldec}, the cells of the quiver Grassmannians attached to $P_{m+1}^{[k]}:=G(\tilde P_{m+1}^{[k]})$ are in one-to-one correspondence with those attached to the lift $\tilde P_{m+1}^{[k]}$.
Moreover, by Theorem~\ref{thm:truncpp}, the quiver Grassmannians $\Gr_{\tbfe}(P_{m+1}^V)$ of arbitrary truncated preprojective representations $P_{m+1}^V$ only depend on $\tbfe$ and $\udim P_{m+1}^V$.
In short, the $2$-quivers $\mathcal Q_{m+1}^{[k]}$ should be viewed as being attached to any (truncated) preprojective representation of the appropriate dimension.

For building the mentioned $2$-quivers, we use the recursive construction described in Section~\ref{Lifting}.
The following remarks are meant to understand the structure of these $2$-quivers. 
\begin{remark}\mbox{}
  \begin{enumerate}
    \item It will turn out that every vertex of the quiver $\mathcal Q_{m+1}^{[k]}$ uniquely corresponds to a vertex of the universal covering quiver $\widetilde{K(n)}$, but that different vertices of $\cQ_{m+1}^{[k]}$ can correspond to the same vertex of $\widetilde{K(n)}$.
      Also every usual arrow of $\mathcal Q_m^{[k]}$ does correspond to an arrow of $\widetilde{K(n)}$ and thus to an arrow of $K(n)$. In particular, the quiver obtained from $\mathcal Q_m^{[k]}$ when forgetting the $2$-arrows gives rise to a (non-disjoint) union of subquivers of $\widetilde{K(n)}$.

@@

\noindent
where the $2$-arrows are colored red.
Note that the vertices $(1,\alpha_i)$ are just the $2$-quivers $\cQ_{1,i}$ corresponding to $\tilde P_{1,i}$.
The $2$-quiver $\mathcal Q_2^{[k]}$ is clearly equivalent to the support quiver (\ref{coeff}) thought of as a 2-quiver with no 2-arrows.
In abuse of notation, we also denote the support quiver by $\mathcal Q_{2}^{[k]}$. In order to keep the illustrations and combinatorics simple, we will work with this quiver instead. We also define the translated two quiver $\cQ_{2,i}$ as those obtained from $\cQ_{2}$ by translating all vertices and ($2$-)arrows by $\alpha_i^{-1}$.
